---
title: "Adaptive reweighting package"
author: "Michael Geuenich"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Adaptive reweighting package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Cel**L** s**E**letion vi**A** a**D**aptiv**E** **R**eweighting (LEADER) is a method to subset cells aiming to capture rare cell types with the goal of creating more balanced datasets. Leader requires a processed seurat object with gene expression data and an optional marker file specifying known marker genes for known cell types.

## Installation
```{r eval=FALSE}
BiocManager::install("leader")
```

## Data processing

The seurat oject must be log normalized and the most variable features must be identified and scaled prior to cell selection with LEADER as follows:

```{r}
library(Seurat)
devtools::load_all()

# Load example data
data(seu, package = "leader")

# Normalize expression and find variable features
seu <- NormalizeData(seu)
seu <- FindVariableFeatures(seu, selection.method = 'vst', nfeatures = 2000)

# Scale gene expression
all.genes <- rownames(seu)
seu <- ScaleData(seu, features = all.genes)
```

Note that the seurat object must also have unique cell ID's as the rownames of the dataframe in the `seu@meta.data` slot. These ID's will then be used to output a subset of cells that can be used as a more balanced dataset.

## Cell selection

Next cells are selected using the LEADER package. This can be in a cell type marker aware way which requires a list of positive and optional negative markers for each cell type expected in the data.

### Cluster based selection

Running LEADER without markers simply requires setting `mode = "cluster_based"`, and determining the required number of cells to select as follows:

```{r}
adaptive_reweighting(seu, n_requested_cells = 40, mode = "cluster_based")
```


### Marker informed selection

The marker informed cell selection requires the creation of a `yaml` file containing marker genes for the expected cell types. These could be positive markers (e.g. CD3 for T-cells) or negative markers (e.g. B220 for T-cells).

The marker yaml file must follow the following structure:

    cell_types:
      CD8 T cells:
        positive:
          - CD3
          - CD8
        negative:
          - B220
          - CD4
      
      CD4 T cells:
        positive:
          - CD3
          - CD4
        negative:
          - CD8
          - B220
          
      B cells:
        positive:
          - B220
          - IgM
      
      ...

The first row must contain the word `cell_types`. Each indented element is an expected cell type with required positive labels and optional negative labels.

To run LEADER using the marker informed selection simply read in a marker file using the `get_markers()` function and add this object to the `adaptive_reweighting` call along with the `mode = "marker_based"` parameter.

```{r}
markers <- get_markers(system.file("extdata", "markers.yml", package = "leader"))
adaptive_reweighting(seu, markers, n_requested_cells = 40, mode = "marker_based")
```

The output of the `adaptive_reweighting` call can then be saved and the cell_id's can be used as more balanced training data.


## Session Information
```{r}
SessionInfo()
```


